Tidal Heating via Internal Gravity Waves in White Dwarf Binaries

### 2019-02-02 writeups/paper.tex

    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
        can be computed. Since the fluid variables have separable as $f(x, z, t)
        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
        \begin{align}
            F_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                    \nonumber\\
                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
                    \label{eq:fpx_lin}
        \end{align}

    \item Finally, it may be noted that horizontally averaged horizontal
        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} =
        0$. Since flux is transported on the group velocity timescale (CITE),
        this simplifies to the known expression for the wave-induced horizontal
        mean flow (cite)
        \begin{equation}
            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
        \end{equation}
        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
        velocity. In terms of our forcing amplitude,
        \begin{equation}
            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
                \label{eq:u0_lin}
        \end{equation}
        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
        \frac{1}{4H^2}}^{3/2}}$.
\end{itemize}
% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
% exact solution of nonlinear.
% Linear solution is globally unstable to RTI though, thus not stable
% equilibrium of IVP
%
% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

@@

The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
in the material derivative, can be solved analytically up to good accuracy. We
may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
must be large enough to be numerically resolved by the spectral code.}. This
system is solved exactly by matching the two homogeneous solutions above and
below $z_0$, We may then approximately relax the solution to $k_{0z}\sigma
\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
compared to the $\delta$-function solution (evaluating the Fourier Transform
of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
\begin{align}
    A(F) ={}& \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}},\\
    u_{1z}(x, z, t) ={}& A(F) \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}

@@

%%%
These are implemented as:
\begin{subequations}\label{se:curr_num}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\Upsilon}{t} - \frac{u_z}{H}
            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                    \cos \p*{k_xx - \omega t},\\
        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
            - \nu \nabla^2 u_x
            &= -\Gamma(z) u_x
                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
                - T_1 \pd{\Upsilon}{x},\\
        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
                - T_1 \pd{\Upsilon}{z},\\
    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
    \end{align}
\end{subequations}
where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
%%%

@@

    \item $F_{px}$ is seen to be constant, as would be expected by
        \autoref{eq:fpx_lin}.

    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
        significantly at higher $z$. This deviation is caused by the top damping
        zone that does not conserve horizontal momentum.
\end{itemize}
We also include a qualitatively similar snapshot of the same simulation run
using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
    \caption{Snapshot of linear simulation after reaching steady state.
        Analytical predictions of $\abs*{u_z}, \bar{U}_0, F_{px}$ are shown in
        orange. Red dotted lines indicate the onset of the damping region and
        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
    used for nonlinear simulations (YUBONOTE should be the same but that
    simulation mysteriously stopped without me noticing; restarting worked, so
    currently rerunning). Note the slight resulting dissipation in $F_{px}$.
    }\label{fig:linear_nu}
\end{figure}

@@

When the fluid absorbs the incident wave, it absorbs the incident horizontal
momentum flux as well, which is converted into additional horizontal momentum of
the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
phase velocity of the incident wave, the critical layer must thus propagate
downwards (towards the wave source) to accommodate the incident momentum. In
other words, the total horizontal momentum of the shear flow obeys conservation
equation
\begin{equation}
    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
        - F_{px} = 0.
\end{equation}
Treating $\bar{U}_0(z > z_c) = \bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
\begin{equation}
    \rho \bar{U}_c\pd{z_c}{t} = -F_{px}.\label{eq:zc_anal}
\end{equation}

@@

Two representative snapshots from our simulation are provided in
\autoref{fig:nl} after the critical layer has had time to form. We may note that
the critical layer, where $F_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
travels downwards as predicted.
\begin{figure}[h]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/nl35_1.png}
        \caption{Earlier time snapshot. Legend is the same as
        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
        third panel.}
    \end{subfigure}

@@

For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
in \autoref{fig:zc_anal}. We see overall good agreement, though some small
deviation is expected since $F_{px}$ is not perfectly conserved owing to
numerical viscosity (and YUBONOTE $F_{px}$ is misestimated?).
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
    simulation.}\label{fig:zc_anal}
\end{figure}

@@

The nonlinear terms in the above equations will transfer energy from lower
wavenumbers to higher wavenumbers. Since spectral codes have no numerical
dissipation, artificial dissipation must be added. To ensure the dissipitive
system conserves horizontal momentum, we begin by adding dissipitive terms to
the flux-conservative form of the Euler fluid equations \autoref{se:fc_orig} (we
use stress tensor $\tau_{ij} = P\delta_{ij}$):
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
            \hat{z} &= 0.
    \end{align}
\end{subequations}
The same $\nu$ is used for both the diffusive and viscous term, though this is
not required. Since the dissipation is not physical and is purely used for
numerical stability, we choose it such that hydrostatic equilibrium is not
modified. Some algebraic manipulation to re-cast it in the form of
\autoref{se:fc_var} gives
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
    \end{align}
\end{subequations}
Hydrostatic equilibrium is still $\vec{\nabla} \varpi_0 + g\hat{z} = 0$ where
$\rho = \rho_0, \vec{u} = 0$. Including the damping layers and forcing terms as
described in \autoref{ss:numerics}, we finally obtain the full system of
equations as simulated in Dedalus:

>>>>

    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
        can be computed. Since the fluid variables have separable as $f(x, z, t)
        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
        \begin{align}
            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                    \nonumber\\
                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
                    \label{eq:fpx_lin}
        \end{align}

    \item Finally, it may be noted that horizontally averaged horizontal
        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
        0$. Since flux is transported on the group velocity timescale (CITE),
        this simplifies to the known expression for the wave-induced horizontal
        mean flow (cite)
        \begin{equation}
            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
        \end{equation}
        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
        velocity. In terms of our forcing amplitude,
        \begin{equation}
            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
                \label{eq:u0_lin}
        \end{equation}
        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
        \frac{1}{4H^2}}^{3/2}}$.
\end{itemize}
% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
% exact solution of nonlinear.
% Linear solution is globally unstable to RTI though, thus not stable
% equilibrium of IVP
%
% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

@@

The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
in the material derivative, can be solved analytically up to good accuracy. We
may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
must be large enough to be numerically resolved by the spectral code.}. This
system is solved exactly by matching the two homogeneous solutions above and
below $z_0$. We may then approximately relax the solution to $k_{0z}\sigma
\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
compared to the $\delta$-function solution (evaluating the Fourier Transform
of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
\begin{align}
    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_zH}}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_zH}}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}

@@

where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
%%%

@@

    \item $S_{px}$ is seen to be constant, as would be expected by
        \autoref{eq:fpx_lin}.

    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
        significantly at higher $z$. This deviation is caused by the top damping
        zone that does not conserve horizontal momentum.
\end{itemize}
We also include a qualitatively similar snapshot of the same simulation run
using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
    \caption{Snapshot of linear simulation after reaching steady state.
        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
        orange. Red dotted lines indicate the onset of the damping region and
        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
    used for nonlinear simulations (YUBONOTE should be the same but that
    simulation mysteriously stopped without me noticing; restarting worked, so
    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
    }\label{fig:linear_nu}
\end{figure}

@@

When the fluid absorbs the incident wave, it absorbs the incident horizontal
momentum flux as well, which is converted into additional horizontal momentum of
the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
phase velocity of the incident wave, the critical layer must thus propagate
downwards (towards the wave source) to accommodate the incident momentum. In
other words, the total horizontal momentum of the shear flow obeys conservation
equation
\begin{equation}
    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}_0(z, t)\;\mathrm{d}z
        - S_{px} = 0.
\end{equation}
Treating $\bar{U}_0(z > z_c) = \bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
\begin{equation}
    \rho \bar{U}_c\pd{z_c}{t} = -S_{px}.\label{eq:zc_anal}
\end{equation}
For constant $S_{px}$ in space and $\rho \approx \rho_0$, this has analytical
solution
\begin{equation}
    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
\end{equation}

@@

Two representative snapshots from our simulation are provided in
\autoref{fig:nl} after the critical layer has had time to form. We may note that
the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
travels downwards as predicted.
\begin{figure}[h]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/nl35_1.png}
        \caption{Earlier time snapshot. Legend is the same as
        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
        third panel.}
    \end{subfigure}

@@

For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
in \autoref{fig:zc_anal}. We see overall good agreement, though some small
deviation is expected since $S_{px}$ is not perfectly conserved owing to
numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
    simulation.}\label{fig:zc_anal}
\end{figure}

@@

The nonlinear terms in the above equations will transfer energy from lower
wavenumbers to higher wavenumbers. Since spectral codes have no numerical
dissipation, artificial dissipation must be added. To ensure the dissipitive
system conserves horizontal momentum, we begin by adding dissipitive terms to
the flux-conservative form of the Euler fluid equations \autoref{se:fc_orig} (we
use stress tensor $\tau_{ij} = P\delta_{ij}$):
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
            \hat{z} &= 0.
    \end{align}
\end{subequations}
The same $\nu$ is used for both the diffusive and viscous term, though this is
not required. Since the dissipation is not physical and is purely used for
numerical stability, we choose it such that hydrostatic equilibrium is not
modified. Some algebraic manipulation to re-cast it in the form of
\autoref{se:fc_var} gives
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
    \end{align}
\end{subequations}
Hydrostatic equilibrium is still $\vec{\nabla} \varpi_0 + g\hat{z} = 0$ where
$\rho = \rho_0, \vec{u} = 0$. Including the damping layers and forcing terms as
described in \autoref{ss:numerics}, we finally obtain the full system of
equations as simulated in Dedalus:
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} ={}& 0,\\
        \partial_t \Upsilon - \frac{u_z}{H}
            ={}& \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
            & - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
                -\Gamma(z) \Upsilon
                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                    \cos \p*{k_xx - \omega t},\\
        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
            \nu \nabla^2 u_x
            - u_x \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
                - \frac{1}{H}\partial_z u_x}
            -\Gamma(z) u_x
                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
                - T_1 \pd{\Upsilon}{x},\\
        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H} ={}&
            \nu \nabla^2 u_z
            - u_z \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
                \frac{1}{H}\partial_z u_{z}}
            -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
            - T_1 \pd{\Upsilon}{z},\\
        \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
            + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
    \end{align}
\end{subequations}
